---
title: "Data 607 Week 3 Assignment"
author: "Sanzida Kaniz"
output: html_document
date: "2024-02-11"
---


```{r}
library(openintro)
library(tidyverse)
```

### 1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"


```{r}

Allages = read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/all-ages.csv", stringsAsFactors = F)

str (Allages)
```

Every major that includes the words "DATA" or "STATISTICS" must be searched for.

```{r}
str_subset(Allages$Major, "DATA|STATISTICS", negate = F)
```

### 2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
Fruits <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")
x<- paste('c(', paste('"',Fruits,'"',sep = "", collapse = ','), sep = "",')')
writeLines(x)

```

### 3 Describe, in words, what these expressions will match:

(.)\1\1
"(.)(.)\\2\\1"
(..)\1
"(.).\\1.\\1"
"(.)(.)(.).*\\3\\2\\1"


"(.)\1\1"
It will search for the first letter at the beginning and check to see if it appears twice more after that.

"(.)(.)\2\1"
I'll examine whether anything matches the opposite by examining the first two starting letters.

"(..)\1"
It will look at the first two letters to check for repetitions.

"(.).\1.\1"
The initial letter—which does not begin on a new line—will be examined first, followed by a character, the first character once more, a character after that, and finally the first character once more.


### 4 Construct regular expressions to match words that:

Start and end with the same character.
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)



```{r}
str_subset(c("civic", "level", "mom", "nose", "eye"), "^(.)((.*\\1$)|\\1?$)")
```


```{r}
str_subset(c("foot", "good", "house", "letter", "data", "church"), "([A-Za-z][A-Za-z]).*\\1")
```

```{r}
str_subset("eleven",  "([a-z]).*\\1.*\\1")
```

